---
title: "Survived Titanic Passenger Prediction with NaiveBayes, Decision Tree, and Random Forest"
author: "Rahma Fairuz Rania"
date: '2022-07-17'
output: 
 html_document:
   toc: true
   toc_float: true
   highlight: zenburn
   df_print: paged
   theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
Case : From data titanic passenger, we want to know does passenger still alive or not. Data can be downloaded here https://www.kaggle.com/c/titanic

# Import Data

```{r}
titanic_train <- read.csv('train.csv')
titanic_test <- read.csv('test.csv')
titanic_gender <- read.csv('gender_submission.csv')
```

# Data Wrangling

To make easier, combine data test and gender into 1 dataframe

```{r, warning=FALSE, message=FALSE}
library(dplyr)
titanic_test <- cbind(titanic_test, Survived = titanic_gender$Survived)
```
- Check missing value

```{r}
# data train
colSums(is.na(titanic_train))
```
There's missing in Age column. Instead of delete it, we change the value with mean of age because we want our model learn more.

```{r}
titanic_train <- titanic_train %>% mutate(Age = if_else(is.na(Age), mean(Age, na.rm = TRUE), Age))

colSums(is.na(titanic_train))
```
```{r}
# data test
colSums(is.na(titanic_test))

titanic_test <- titanic_test %>% na.omit()
anyNA(titanic_test)
```
- Check structure of data

```{r}
str(titanic_train)
```

Here are some information about columns in wholesale data
* `PassengerId`: Id number of passenger
* `Survived` : Passenger Survival or Not
* `Pclass` :  A proxy for socio-economic status (SES). 1st = Upper, 2nd = Middle, 3rd = Lower
* `Name` : Name of Passenger
* `Sex` : Sex of Passenger 
* `Age` : Age of Passenger
* `SibSp` : Sibling = brother, sister, stepbrother, stepsister; Spouse = husband, wife (mistresses and fiancés were ignored)
* `Parch` :  Parent = mother, father; Child = daughter, son, stepdaughter, stepson; Some children travelled only with a nanny, therefore parch=0 for them.
* `Ticket` : Ticket number
* `Fare` : Passenger fare
* `Cabin` : Cabin number
* `Embarked` : Port of Embarkation 

- Change column into right data type and delete some column

```{r}
titanic_train <- titanic_train %>% select(-c(PassengerId, Name, Ticket, Cabin)) %>% 
  mutate(Survived = as.factor(Survived),
         Pclass = as.factor(Pclass),
         Sex = as.factor(Sex),
         SibSp = as.factor(SibSp),
         Parch = as.factor(Parch),
         Embarked = as.factor(Embarked))

titanic_test <- titanic_test %>% select(-c(PassengerId, Name, Ticket, Cabin)) %>% 
  mutate(Survived = as.factor(Survived),
         Pclass = as.factor(Pclass),
         Sex = as.factor(Sex),
         SibSp = as.factor(SibSp),
         Parch = as.factor(Parch),
         Embarked = as.factor(Embarked))
```

# Data Pre-processing

- Check class proportion

```{r}
prop.table(table(titanic_train$Survived))
```
Our data still balance but we can make balance proportion with downsampling

```{r, warning=FALSE, message=FALSE}
RNGkind(sample.kind = "Rounding")
set.seed(555)
library(caret)
titanic_train <- downSample(x = titanic_train %>% 
                           select(-Survived),
                          y = titanic_train$Survived,
                         yname = "Survived") 
```


```{r}
prop.table(table(titanic_train$Survived))
```
Now our data is balance.

# Naive Bayes Model

- Build model with `naiveBayes()` function from library `e1071`

```{r, warning=FALSE, message=FALSE}
library(e1071)
titanic_nb <- naiveBayes(Survived ~., titanic_train, laplace = 1)
titanic_nb
```

```{r}
prop.table(table(titanic_train$Survived, titanic_train$Sex), margin = 2)
```
> From the class proportion above, female passenger that survive is 81% meanwhile male passenger 27%. Female passenger prioritized.

- Model evaluation

```{r}
titanic_test$pred <- predict(titanic_nb, titanic_test, type = 'class')
```

- Evaluation with confusionMatrix from library caret

```{r}
library(caret)
confusionMatrix(titanic_test$pred, reference = titanic_test$Survived)
```
- ROC model naiveBayes

ROC (Receiver Operating Curve) describe relation between True Positif Rate(TPR) or Recall/Sensitivity and False Positif Rate(FPR) or (1-specificity). Ideal model has TPR > FPR
Build dataframe ROC, assume positive class is 1(survived).

```{r}
titanic_test$pred <- predict(titanic_nb, newdata = titanic_test, type = "raw")
titanic_test$actual <- ifelse(titanic_test$Survived == 1, 'Alive', 'Death')
```

Build ROC 

```{r, warning=FALSE, message=FALSE}
library(ROCR)

# objek prediction
roc_pred <- prediction(predictions = titanic_test$pred[,1], labels = titanic_test$actual)

# ROC curve
plot(performance(prediction.obj = roc_pred, measure = "tpr", x.measure = "fpr"))
abline(0,1,lty = 8)
```

- AUC model naivebayes

After we search for ROC, next we search value under ROC plot with AUC(Area Under Curve).

```{r}
titanic_auc <- performance(prediction.obj = roc_pred, measure = 'auc')
titanic_auc@y.values
```
our Naive Bayes model is good enough to separate positive(alive) and negative(dead) class.

# Decision Tree

- Check class proportion 

```{r}
prop.table(table(titanic_train$Survived))
```
- Build model with `ctree()` from library `partykit`

```{r, warning=FALSE, message=FALSE}
library(partykit)
titanic_tree <- ctree(Survived ~., titanic_train)
titanic_tree
```
For better information, visualize it

```{r}
plot(titanic_tree, type = 'simple')
```
From tree structure above, we can see classification is based on `sex`, `Pclass`, `Fare`, and `Age`. Information of tree :
```{r}

```
- Root node (highest level of tree) : Sex, means that gender rescuers consider gender first. This node is very important to determine target.
- Interior node : Pclass, Fare, Age, SibSp. Second branch used if root is not enough to determine target.
- Leaf node : [3], [5], [6], [8], [11], [12], [13]. Predict target value.


- Model evaluation

```{r}
# predict data test
titanic_test_pred <- predict(object = titanic_tree, newdata = titanic_test, type = "response")

# confusion matrix data test
confusionMatrix(data = titanic_test_pred, reference = titanic_test$Survived, positive = "1")
```

# Random Forest

For speed up computation, we can delete non-informative variable.

```{r}
# check non-informative var
nearZeroVar(titanic_train)
```
There's no non-informative so we can use Random Forest for building model.

- Build model with k-fold validation 

```{r, warning=FALSE, message=FALSE}
library(randomForest)
set.seed(572)
ctrl <- trainControl(method = "repeatedcv",
                    number = 5, # k-fold
                     repeats = 3) # repeat
titanic_forest <- train(Survived ~., data = titanic_train, method = 'rf', trControl = ctrl)
titanic_forest
```
Optimum k splitting at each tree node is 11. 

- Check the important variable using varImp()

```{r}
varImp(titanic_forest)
```

```{r}
titanic_forest$finalModel
```

- Predict random forest model

```{r}
titanic_forest_pred <- predict(titanic_forest, newdata = titanic_test, type = "raw")
```

- Model evaluation

```{r}
confusionMatrix(data = titanic_forest_pred, reference = titanic_test$Survived, positive = "1")
```

# Conclusion

Naive Bayes model has
```{r}

```

- 81% accuracy
- 84% sensitivity
- 75% specificity
- 84% precision


Decision Tree model has
```{r}

```

- 81% accuracy
- 98% sensitivity
- 71% specificity
- 67% precision

Random Forest model has
```{r}

```

- 81% accuracy
- 85% sensitivity
- 78% specificity
- 71% precision

> In this case, we focused on Sensitivity. We want model can predict positive class from the actual is positive. So, Decision Tree is the best model which has highest sensitivity.